Is Archaeology a science? Insights and imperatives from 10,000 articles and a year of reproducibility reviews

### 2025-04-19 analysis/paper/paper.qmd
polish the figures

base_size <- 6
color <- c('#1b9e77','#d95f02','#7570b3')

# Number of authors ------------------
boxlplot_n_authors <- 
  ggplot() +
  # boxplot of data from this study
  geom_boxplot(data = items_df %>% 
  filter(!is.na(year)),
  aes(1, log(authors_n)),
               size = 1)  +
  # boxplot of data from Fanelli & Glänzel Fig 2
  geom_boxplot(data = sim_data %>% 
  filter(Variable == "N. of authors (ln)",
         Category %in% c("h", "p", "s")),
  aes(1, Value, 
      group = Category),
  size = 1,
  colour = color, 
  alpha = 0)  +
  # annotations from Fanelli & Glänzel Fig 2
  geom_text_repel(data = sim_data %>%  
    filter(Variable == "N. of authors (ln)",
           Category %in% c("h", "p", "s")) %>%  
    group_by(Category) %>%  
    summarise(y = median(Value)) %>%
    mutate(
      label = as.character(Category)
    ),
                  aes(c(0.75, 1, 1.25), y, label = label),
                  color = color,
                  bg.colour = "white", 
                  bg.r = .2, 
                  force = 0) +
  scale_y_continuous(limits = c(0, 5)) +
  scale_x_continuous(labels = NULL) +
  theme_minimal(base_size = base_size) + 
  theme(panel.grid  = element_blank()) +
  ylab("N. of authors (ln)") +
  xlab("Collaborator group size") 
```

```{r}
# Relative title length  ----------------

@@

boxlplot_rel_title_length <- 
items_df_title %>% 
  filter(!is.na(year)) %>% 
  ggplot(aes(1,
             relative_title_length)) +
  geom_boxplot(
               size = 1)  +
  # boxplot of data from Fanelli & Glänzel Fig 2
  geom_boxplot(data = sim_data %>% 
  filter(Variable == "Relative title length (ln)",
         Category %in% c("h", "p", "s")),
  aes(1, Value, 
      group = Category),
  size = 1,
  colour = color, 
  alpha = 0)  +
  # annotations from Fanelli & Glänzel Fig 2
  geom_text_repel(data = sim_data %>%  
    filter(Variable == "Relative title length (ln)", 
           Category %in% c("h", "p", "s")) %>%  
    group_by(Category) %>%  
    summarise(y = median(Value)) %>%
    mutate(
      label = as.character(Category)
    ),
                  aes(c(0.75, 1, 1.25), y, label = label),
                  bg.colour = "white", 
    colour = color,
                  bg.r = .2, 
                  force = 0) +
  scale_y_continuous(limits = c(-4.5, 3),
                     breaks =  seq(-5, 5, 1),
                     labels = seq(-5, 5, 1)) +
  scale_x_continuous(labels = NULL) +
  theme_minimal(base_size = base_size) +
  theme(panel.grid  = element_blank()) +
  ylab("Ratio of title length to article length (ln)") +
  xlab("Relative title length")

# Number of pages ------------------
boxlplot_n_pages <- 
items_df %>% 
  ggplot(aes(1,
             log(pages_n))) +
  geom_boxplot(
               size = 1)  +
  # boxplot of data from Fanelli & Glänzel Fig 2
  geom_boxplot(data = sim_data %>% 
  filter(Variable == "N. of pages (ln)",
         Category %in% c("h", "p", "s")),
  aes(1, Value, 
      group = Category),
  size = 1,
  colour = color, 
  alpha = 0)  +
  # annotations from Fanelli & Glänzel Fig 2
  geom_text_repel(data = sim_data %>%  
    filter(Variable == "N. of pages (ln)", 
           Category %in% c("h", "p", "s")) %>%  
    group_by(Category) %>%  
    summarise(y = median(Value)) %>%
    mutate(
      label = as.character(Category)
    ),
                  aes(c(0.75, 1, 1.25), y, label = label),
                  bg.colour = "white",
    colour = color, 
                  bg.r = .2, 
                  force = 0) +
  scale_y_reverse(limits = c(5, 0)) +
  scale_x_continuous(labels = NULL) +
  theme_minimal(base_size = base_size) + 
  theme(panel.grid  = element_blank()) +
  ylab("N. of pages (ln)") +
  xlab("Article length")

@@

# plot
boxlplot_price_index <- 
items_df %>% 
  ggplot(aes(1,
             prices_index)) +
  geom_boxplot(
               size = 1)  +
  # boxplot of data from Fanelli & Glänzel Fig 2
  geom_boxplot(data = sim_data %>% 
  filter(Variable == "Price's index",
         Category %in% c("h", "p", "s")),
  aes(1, Value, 
      group = Category),
  size = 1,
  colour = color,  
  alpha = 0)  +
  # annotations from Fanelli & Glänzel Fig 2
  geom_text_repel(data = sim_data %>%  
    filter(Variable == "Price's index", 
           Category %in% c("h", "p", "s")) %>%  
    group_by(Category) %>%  
    summarise(y = median(Value)) %>%
    mutate(
      label = as.character(Category)
    ),
                  aes(c(0.75, 1, 1.25), y, label = label),
                  bg.colour = "white", colour = color, 
                  bg.r = .2, 
                  force = 0) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_continuous(labels = NULL) +
  theme_minimal(base_size = base_size) +
  theme(panel.grid  = element_blank()) +
  ylab("Prop. refs in last 5 years") +
  xlab("Recency of references")

@@

# plot
boxlplot_shannon_index <- 
shannon_per_item %>% 
  filter(!is.na(year)) %>% 
  ggplot(aes(1,
             shannon)) +
  geom_boxplot(
               size = 1)  +
   # boxplot of data from Fanelli & Glänzel Fig 2
  geom_boxplot(data = sim_data %>% 
  filter(Variable == "Shannon div. of sources",
         Category %in% c("h", "p", "s")),
  aes(1, Value, 
      group = Category),
  size = 1,
  colour = color,  
  alpha = 0)  +
  # annotations from Fanelli & Glänzel Fig 2
  geom_text_repel(data = sim_data %>%  
    filter(Variable == "Shannon div. of sources", 
           Category %in% c("h", "p", "s")) %>%  
    group_by(Category) %>%  
    summarise(y = median(Value)) %>%
    mutate(
      label = as.character(Category)
    ),
                  aes(c(0.75, 1, 1.25), y, label = label),
    colour = color, 
                  bg.colour = "white", 
                  bg.r = .2, 
                  force = 0) +
  scale_y_reverse(limits = c(6, 0)) +
  scale_x_continuous(labels = NULL) +
  theme_minimal(base_size = base_size)  +
  theme(panel.grid  = element_blank()) +
  ylab("Shannon Index") +
  xlab("Diversity of references")
```

>>>>

base_size <- 6
color <- c('#d95f02', '#7570b3',  '#1b9e77')
alpha <- 0.2
linewidth <- 0.1

# Number of authors ------------------
boxlplot_n_authors <- 
  ggplot() +
  # boxplot of data from this study
  geom_boxplot(data = items_df %>% 
  filter(!is.na(year)),
  aes(1, log(authors_n)),
               size = 1)  +
  # boxplot of data from Fanelli & Glänzel Fig 2
  geom_boxplot(data = sim_data %>% 
  filter(Variable == "N. of authors (ln)",
         Category %in% c("h", "p", "s")),
  aes(1, Value, 
      group = Category),
  size = 1,
  fill = color,
  colour = color, 
  alpha = alpha,
  linewidth = linewidth)  +
  # annotations from Fanelli & Glänzel Fig 2
  geom_text_repel(data = sim_data %>%  
    filter(Variable == "N. of authors (ln)",
           Category %in% c("h", "p", "s")) %>%  
    group_by(Category) %>%  
    summarise(y = median(Value)) %>%
    mutate(
      label = as.character(Category)
    ),
                  aes(c(0.75, 1, 1.25), y, label = label),
                  color = color,
                  bg.colour = "white", 
                  bg.r = .2, 
                  force = 0) +
  scale_y_continuous(limits = c(0, 5)) +
  scale_x_continuous(labels = NULL) +
  theme_minimal(base_size = base_size) + 
  theme(panel.grid  = element_blank()) +
  ylab("N. of authors (ln)") +
  xlab("Collaborator group size")

# Relative title length  ----------------

@@

boxlplot_rel_title_length <- 
items_df_title %>% 
  filter(!is.na(year)) %>% 
  ggplot(aes(1,
             relative_title_length)) +
  geom_boxplot(
               size = 1)  +
  # boxplot of data from Fanelli & Glänzel Fig 2
  geom_boxplot(data = sim_data %>% 
  filter(Variable == "Relative title length (ln)",
         Category %in% c("h", "p", "s")),
  aes(1, Value, 
      group = Category),
  size = 1,
  fill = color,
  colour = color, 
  alpha = alpha,
  linewidth = linewidth)  +
  # annotations from Fanelli & Glänzel Fig 2
  geom_text_repel(data = sim_data %>%  
    filter(Variable == "Relative title length (ln)", 
           Category %in% c("h", "p", "s")) %>%  
    group_by(Category) %>%  
    summarise(y = median(Value)) %>%
    mutate(
      label = as.character(Category)
    ),
                  aes(c(0.75, 1, 1.25), y, label = label),
                  bg.colour = "white", 
    colour = color,
                  bg.r = .2, 
                  force = 0) +
  scale_y_continuous(limits = c(-4.5, 3),
                     breaks =  seq(-5, 5, 1),
                     labels = seq(-5, 5, 1)) +
  scale_x_continuous(labels = NULL) +
  theme_minimal(base_size = base_size) +
  theme(panel.grid  = element_blank()) +
  ylab("Ratio of title length to article length (ln)") +
  xlab("Relative title length")

# Number of pages ------------------
boxlplot_n_pages <- 
items_df %>% 
  ggplot(aes(1,
             log(pages_n))) +
  geom_boxplot(
               size = 1)  +
  # boxplot of data from Fanelli & Glänzel Fig 2
  geom_boxplot(data = sim_data %>% 
  filter(Variable == "N. of pages (ln)",
         Category %in% c("h", "p", "s")),
  aes(1, Value, 
      group = Category),
  size = 1,
  fill = color,
  colour = color, 
  alpha = alpha,
  linewidth = linewidth)  +
  # annotations from Fanelli & Glänzel Fig 2
  geom_text_repel(data = sim_data %>%  
    filter(Variable == "N. of pages (ln)", 
           Category %in% c("h", "p", "s")) %>%  
    group_by(Category) %>%  
    summarise(y = median(Value)) %>%
    mutate(
      label = as.character(Category)
    ),
                  aes(c(0.75, 1, 1.25), y, label = label),
                  bg.colour = "white",
    colour = color, 
                  bg.r = .2, 
                  force = 0) +
  scale_y_reverse(limits = c(5, 0)) +
  scale_x_continuous(labels = NULL) +
  theme_minimal(base_size = base_size) + 
  theme(panel.grid  = element_blank()) +
  ylab("N. of pages (ln)") +
  xlab("Article length")

@@

# plot
boxlplot_price_index <- 
items_df %>% 
  ggplot(aes(1,
             prices_index)) +
  geom_boxplot(
               size = 1)  +
  # boxplot of data from Fanelli & Glänzel Fig 2
  geom_boxplot(data = sim_data %>% 
  filter(Variable == "Price's index",
         Category %in% c("h", "p", "s")),
  aes(1, Value, 
      group = Category),
  size = 1,
  fill = color,
  colour = color,  
  alpha = alpha,
  linewidth = linewidth)  +
  # annotations from Fanelli & Glänzel Fig 2
  geom_text_repel(data = sim_data %>%  
    filter(Variable == "Price's index", 
           Category %in% c("h", "p", "s")) %>%  
    group_by(Category) %>%  
    summarise(y = median(Value)) %>%
    mutate(
      label = as.character(Category)
    ),
                  aes(c(0.75, 1, 1.25), y, label = label),
                  bg.colour = "white", colour = color, 
                  bg.r = .2, 
                  force = 0) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_continuous(labels = NULL) +
  theme_minimal(base_size = base_size) +
  theme(panel.grid  = element_blank()) +
  ylab("Prop. refs in last 5 years") +
  xlab("Recency of references")

@@

# plot
boxlplot_shannon_index <- 
shannon_per_item %>% 
  filter(!is.na(year)) %>% 
  ggplot(aes(1,
             shannon)) +
  geom_boxplot(
               size = 1)  +
   # boxplot of data from Fanelli & Glänzel Fig 2
  geom_boxplot(data = sim_data %>% 
  filter(Variable == "Shannon div. of sources",
         Category %in% c("h", "p", "s")),
  aes(1, Value, 
      group = Category),
  size = 1,
  fill = color,
  colour = color,  
  alpha = alpha,
  linewidth = linewidth)  +
  # annotations from Fanelli & Glänzel Fig 2
  geom_text_repel(data = sim_data %>%  
    filter(Variable == "Shannon div. of sources", 
           Category %in% c("h", "p", "s")) %>%  
    group_by(Category) %>%  
    summarise(y = median(Value)) %>%
    mutate(
      label = as.character(Category)
    ),
                  aes(c(0.75, 1, 1.25), y, label = label),
    colour = color, 
                  bg.colour = "white", 
                  bg.r = .2, 
                  force = 0) +
  scale_y_reverse(limits = c(6, 0)) +
  scale_x_continuous(labels = NULL) +
  theme_minimal(base_size = base_size)  +
  theme(panel.grid  = element_blank()) +
  ylab("Shannon Index") +
  xlab("Diversity of references")
```
